---
title: "Pratical Machine Learning"
author: "Filippo Quarta"
date: "Monday, November 23, 2015"
output: html_document
---

This code has ben done as an assignment for the Practical Machine Learning course on Coursera. 


**Getting Data**

Data are retrieved from the web source and loaded into memory as data frames.

```{r}
library(caret)


#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",destfile = "pml-training.csv")
#download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")

training <- read.csv("pml-training.csv")
test <- read.csv("pml-testing.csv")

```

**Data cleaning and partitioning data**

Only variables related to sensors measurements are maintened and the original training set is divided into training and validation test ( equally sized ). Columns containing more than 80% of NA's are removed.

```{r}

training <- training[,7:160]
test <- test[,7:160]

vars <- colSums(is.na(test))<0.8*nrow(test)
training <- training[,vars]
test <- test[,vars]

index <- createDataPartition(training$classe, p = 0.5, list=FALSE)
trains <- training[index,]
valids <- training[-index,]

```

**Training the model**

A random forest algorithm is applied using 4-fold cross validation for the train set.

```{r}

modFit <- train(classe ~ .,data=trains,method="rf",
                          trControl=trainControl(method="cv",number=4),
                          prox=TRUE,allowParallel=TRUE)

print(modFit$finalModel)
```


**Validated the model**

The validation set is used in order to verify the accuracy of the model.

```{r}

confusionMatrix(valids$classe, predict(modFit, valids))

```

